### 2021-05-04 AST-Framing-Ontario/Data Pre-Processesing.rmd
May 4th update

---
title: "Data Pre-Processing"
author: "Jason Lam"
date: "11/03/2021"
output:
  pdf_document: default
  word_document: default
---

# Introduction

This notebook provides a methodology for converting portable document formats (PDFs) into a text corpus that can be used in natural language processing (NPL) analysis. The text corpus is a collection of structured texts which enables analysis using packages such as `topicmodels`. This notebook serves as an introduction on how to convert files and conduct a surface level analysis.

Instructions are provided below for processing a single folder of PDF documents, as a step-by-step guide of the processing techniques used. Another set of instructions is provided at the end of this notebook for the processing multiple folders simultaneously.

# Preliminaries

Clear the environment: 
```{r}
rm(list = ls())
```

Load the following packages:
```{r}
library(tm)
library(readr)
library(tools)
library(dplyr)
library(tidytext)
library(topicmodels)
library(ggplot2)
library(wordcloud)
#library(word2vec)
library(pdftools)
library(reshape2)
library(text2vec)
```

*Parametrize the document to run different corpus* 
[We could parametrize specific stop words base on the specific files]

# Accessing Data: Grey Literature

Active school travel (AST) and school travel planning (STP) initiatives typically bring together a group of stakeholders from the planning, transportation, public health, and education sectors. These stakeholders may or may not publish information about their involvement in AST and STP efforts on their respective websites or in policy documents.

First, we do a preliminary scan of current AST and/or STP projects funded by the Ontario Active School Travel fund to determine which stakeholders are involved in such initiatives. We find that AST projects are often led by a school board in collaboration with transportation planners, public health professionals, and regional transportation consortia. Non-profit organizations, police services, and advocacy groups are other common stakeholders who play a role in support AST and STP.

Next, we create one folder for each stakeholder group involved in AST and/or STP efforts. The folders below were used in the framing analysis; they are provided in this notebook as an example to recreate, modify, or extend our analysis.

For these three following folders, webpages were sourced directly from stakeholder websites from three main groups: transportation consortia, municipalities or regions, and school boards. These webpages were saved as PDF documents via the web browser.

Inclusion of webpages was determined based on the following considerations:
1) It was important that webpages were easy to find as the analysis pertains to how such issues are framed to the public. Our interpretation of 'easy to find' are webpages that required no more than 2-4 separate links from the initial browser search. 
2) The analysis focuses on the province of Ontario, not a single municipality or region. For this reason, the search was guided by a list of all school boards across the province. The websites of each school board were manually searched to determine whether they discussed AST and/or STP. 
3) Consortia and municipalities/regions were organized based on geographic area (i.e., proximity to the school boards).

__Transportation Consortia__

Transportation consortia are dedicated transportation bodies that primarily focus on providing the school busing service to families in their associated region.

In total, 9 webpages were pulled from 7 of the 39 consortia examined in Ontario. Webpages are determined to be relevant if they specifically pertain and/or mention AST or STP.

__Municipalities or Regions__

Municipalities or regions have webpages about transportation plans, programs, and services provided, some of which are not relevant to our analysis. Only webpages specifically discussing active travel, AST, or STP were included. Webpages were also sourced from associated public health units because they have a direct relation to governing bodies at the municipal or regional level.

In total, 29 webpages were pulled.

__School Boards__

These bodies use funding provided to them by the province to run a collection of schools. They prepare the annual budget, hire teachers and staff, supervise the education program, and other tasks to ensure schools are following all regulations set by the Province of Ontario's *Education Act*. Only English speaking public and Catholic school boards were considered.

Pages mentioning STP and/or AST were prioritized. However, if they did not exist, pages discussing transportation were downloaded instead. In total, 32 webpages were pulled from 31 of the 61 school boards examined in Ontario. Of the remaining the were not included, 22 linked their respective transportation consortia, 7 did not have pages that meet the criteria, 1 met the criteria but could not be downloaded in a PDF format.

# Accessing Data: Research Literature

__Academic Papers__

A literature search was conducted in _Web of Science_ using "active", "travel", "school" as keywords, and then further narrowed down by discipline "geography", "transportation, "planning".

In total, 233 articles were included from a list of 253. Reasons of exclusions for remaining 20 include unavailablity online, access only on print, and published in non-english.

# Conversion Process

The conversion process for this section can only be conducted on one file at a time. We select one of the files provided to use in this package as an example of the process. Alternatively, you can use a file relevant to your municipality/region. To run multiple files or to create separate corpora, refer to the section below.

## Searchable PDF files to .txt conversion

PDF documents are brought into the corpus directly, however each page of the document is read as an individual document. For our purposes, a conversion from PDF to .txt will allow us to keep all pages in a document together when loading it into the corpus. Make sure you have the directory path to where your documents are stored and re-assign the `pdfLocation` to that corresponding location. R uses '/' as the directory separator.

```{r}
pdfLocation <- "Texts/Academic Papers NoRef"

# Creates folder for .txt files
if (file.exists(paste(pdfLocation, "txt"))){
    txtLocation <- paste(pdfLocation, "txt")
  }else{
    dir.create(paste(pdfLocation, "txt"))
    txtLocation <- paste(pdfLocation, "txt")
  }

# Creates a complete list of full directory paths and file names (with extension .pdf) from the pdf location as specified above
fileBaseNames <- list.files(pdfLocation, pattern = "*.pdf", full.names=FALSE)
filePath <- list.files(pdfLocation, pattern = "*.pdf", full.names=TRUE)
n <- 1

for (address in filePath)
{
doc <- readPDF(control = list(text = "-layout"))(elem = list(uri = address), language = "en")
test <- doc$content

# `file_path_sans_ext(basename())` from the tools package strips the base name for the file of interest and removed the extension at the end which as seen below allow for the replacement of .pdf for .txt

# use of `sep = ""` argument is important to prevent the addition of spaces, which is the default for the paste function 
newFile <- paste(txtLocation, "/", file_path_sans_ext(basename(address)), ".txt", sep = "")
write(test, file = newFile)
print( paste(file_path_sans_ext(basename(address)), ".txt Has been created", sep = ""))

# Creates uniquely id text store in 'Environment'
assign(paste("text", n, sep = ""), readChar(newFile, file.info(newFile)$size))
n <- n+1
}
```

## Creating data frame of texts

Creation of a data frame of texts is required prior to conversion to corpus. The data frame must contain 'doc_id' and 'text'. Additional fields may be included in the data frame, however it will be treated as metadata by the corpus.

```{r}
# This loop creates a vector of the texts
text <- vector()
doc_id <- c(2:n-1)

for (i in seq(1, n-1, 1))
{
text[i] <- eval(parse(text = paste("text", i, sep = "")))
}

# Turns vector into data frame 
textDF <- data.frame(doc_id, text)
```

## Conversion of data frame to corpus

Next, convert the data frame to a corpus:
```{r}
df_source <- DataframeSource(textDF)
corpus <- VCorpus(df_source)
# Example 
content(corpus[[1]])
```

## Corpus cleaning

Many words should be removed while cleaning the corpus to ensure that the analysis does not pick up on common nouns, adjectives, or indefinite articles. Words can be removed manually by creating a vector of strings. Add strings to 'NAwords': 
```{r}
corpus <- tm_map(corpus, stripWhitespace)
corpus <- tm_map(corpus, content_transformer(tolower))

NAwords <- c("school", "transportation", "can", "use", "active", "et", "al", "travel", "study", "elsevier", "na", "sciencedirect", "journal", "transport", "mode", "modes", "children", "the", "use", "n", "s", "schools", "also", "p", "b", "used", "number")
corpus <- tm_map(corpus, removeWords, NAwords)
corpus <- tm_map(corpus, removeNumbers)

corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeWords, stopwords("english"))

# Example
content(corpus[[1]])
```

## Term Document Matrix & Document Term Matrix

[Insert brief description of this step]

```{r}
TDM <- TermDocumentMatrix(corpus)
DTM <- DocumentTermMatrix(corpus)
matrix1 <- as.matrix(TDM)
matrix2 <- as.matrix(DTM)

```

## Natural Languange Processing (NLP) Analysis

NLP is a way to correlate words across a collection of texts or corpus.

### Topic Modelling with Latent Dirichlet Allocation (LDA)

The goal of topic modelling is to identify the re-occurring themes in our corpus quantitatively with code. The code below utilizes LDA, an algorithm that when provided a number of discrete topics, `k`, will calculate the likelihood of any given term being generated in a given topic. This algorithm runs under the 'bag of words' assumptions with means grammatical structure is not considered. Only words are rather considered.

```{r}
# k represents the number of topics to be generated by the algorithm 
k <- 4
lda <- LDA(DTM, k, control = list(seed = 1234))

topics <- tidy(lda, matrix = "beta")

# The value below determines the number of top terms
top_terms <- topics %>% 
  group_by(topic) %>% 
  top_n(15, beta) %>% 
  ungroup() %>% 
  arrange(topic, -beta)
top_terms

# Beta represents the probability of 'term' being generated in topic 'n' 
# Topics are generated by the LDA algorithm, its is up to the user to interpret the context of the topics 
terms(lda,15)
top_terms %>% 
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot(aes(beta, term, fill = factor(topic))) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~topic, scales = "free") + 
  scale_y_reordered() +
  scale_x_continuous(limits = c(0,0.06))

```

### LDA using multiple seeds

This is a more robust approach since LDA has some random components on how it begins developing its clusters. By taking the average of multiple iterations with random seeds, we .... [_finish sentence_]

NOTE: RUNNING THE BLOCK BELOW WITH 'Academic Articles'AS INPUT WILL RESULT IN VERY LONG PROCESSING TIME

```{r}
# 
# # Generate a random seed to use 
# numIteration <- 10
# r <- 1 
# col <- 4 
# betaList <- c("beta.0")
# seeds <- sample.int(10000, numIteration-1)
# avgtopics <- topics
# 
# for (seed1 in seeds){
#   templda <- LDA(DTM, k, control = list(seed = seed1))
#   temptopics <- tidy(templda, matrix = "beta")
#   avgtopics <- left_join(avgtopics, temptopics , by = c("topic", "term"))
#   colnames(avgtopics)[col] <- paste("beta.", r, sep = "")
#   betaList <- append(betaList, paste("beta.", r, sep = ""))
#   r <- r + 1
#   col <- col + 1 
# }
# colnames(avgtopics)[3] <- "beta.0"
# 
# avgtopics$beta.mean <- rowMeans(avgtopics[betaList])
# 
# # The value below determines the number of top terms
# avg_top_terms <- avgtopics %>% 
#   group_by(topic) %>% 
#   top_n(15, beta.mean) %>% 
#   ungroup() %>% 
#   arrange(topic, -beta.mean)
# avg_top_terms
# 
# # Beta represents the probability of 'term' being generated in topic 'n' 
# # Topics are generated by the LDA algorithm, it's up to the user to interpret the context of the topics 
# 
# avg_top_terms %>% 
#   mutate(term = reorder_within(term, beta.mean, topic)) %>%
#   ggplot(aes(beta.mean, term, fill = factor(topic))) + 
#   geom_col(show.legend = FALSE) + 
#   facet_wrap(~topic, scales = "free") + 
#   scale_y_reordered() +
#   scale_x_continuous(limits = c(0,0.06))

```

## Tokenization by n-gram

Here we examine texts based on the structural relationship of a sequence of words. We can "tokenize" n-grams, where `n` represents the number of consecutive words, to chart what group of words most frequently occur in our corpus.

```{r}
# Converting corpus to tidytext data type 
td_corpus <- tidy(corpus)

# n can be changed 
test_bigrams <- td_corpus %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2)

test_trigrams <- td_corpus %>%
  unnest_tokens(trigram, text, token = "ngrams", n = 3)
# Bigrams & Trigrams that are most common

top_bigrams <- test_bigrams %>%
  count(bigram, sort = TRUE)

top_trigrams <- test_trigrams %>% 
  count(trigram, sort = TRUE)

top_bigrams
top_trigrams
```

## Using TDM to make a word cloud

Word cloud is a nice visual way to view the most common words in your text document matrix

```{r}
v <- sort(rowSums(matrix1), decreasing= TRUE)
d <- data.frame(word = names(v), freq=v)
set.seed(4321)
wordcloud(d$word, d$freq, min.freq = 10, max.words = 50, random.order = FALSE, rot.per = 0.5, random.color = TRUE)
```

# GloVe Word Embeddings

https://m-clark.github.io/text-analysis-with-R/word-embeddings.html

```{r}
# Create new data frame with the cleaned corpus 
cleanTextDF <- data.frame(text = sapply(corpus, as.character), stringsAsFactors = FALSE)

it_train = itoken(cleanTextDF$text, progressbar = FALSE)
vocab = create_vocabulary(it_train)

# Cut out terms that have a minimum count of 5 
vocab <- prune_vocabulary(vocab, term_count_min = 5)
vocab

vectorizer <- vocab_vectorizer(vocab)
tcm <- create_tcm(it_train, vectorizer, skip_grams_window = 10)

glove = GlobalVectors$new(rank = 50, x_max = 10)
main = glove$fit_transform(tcm, n_iter = 10, convergence_tol = 0.001)
context = glove$components

word_vectors = main + t(context)

safe = word_vectors["wellbeing", , drop = FALSE] - word_vectors["physical", , drop = FALSE] + word_vectors["mental", , drop = FALSE]

cos_sim = sim2(x = word_vectors, y = safe, method = "cosine")

head(sort(cos_sim[,1], decreasing = TRUE), 10)
```

# Multi-Files

Clear Environment 
```{r}
rm(list = ls())
```

`masterFolder` will be the location of your sub-folders, containing your PDFs. Unlike in 'PDF to Corpus.Rmd' you do not need to create .txt folder corresponding to each PDF folder, this is done automatically by the code.

IMPORTANT: The names of the folders cannot contain the consecutive characters `txt`, the code filters out those folders. 
NOTE: The order of files below will also be the order the corpus are created, the index will match the corpus number 
```{r}
masterFolder <- "Texts"

subFolders <- list.files(masterFolder, full.names=TRUE)
# if txt folder already exists from previous execution of code this ensures it is not added to subfolders 
# The limitation is that your file name, at any point cannot contain the sequence of characters "txt" as it will be flagged as a non-pdf subfolder

subFolders <- subFolders[grep(x =list.files(masterFolder, full.names=TRUE), pattern =  "txt", invert = TRUE)]
subFolders
```

Corpora will be cleaned for white spaces, punctuation, English stop words, and converted to lowercase. Any additional stop words will need to be removed manually as the user sees fit. 'corpusKey' is a data frame that indicates which file a corpus was derived from. 
```{r}
# Stop words for each of the corpora, or based on how your directory organizes files 
# Add stop words if you have more than 4 corpora 
NAwords2 <- c("login", "css", "sir", "waterloo", "niagara", "john", "school", "transportation", "york", "ottawa", "can", "tip", "london", "provides", "elgin", "find", "make", "active", "travel", "lord", "holy", "toronto", "-")
NAwords3 <- c("york", "lane", "school", "transportation", "region", "halton", "can", "use", "guelph", "kingston", "durham", "get")
NAwords4 <- c("active", "and", "asst", "board", "can", "cancellations", "catholic", "contact", "cycling", "district", "east", "gta", "https", "login", "may", "or", "please", "school", "schools", "stp", "student", "students", "the", "transportation", "tri", "walking", "waterloo", "website", "will")
NAwords1 <- c("school", "transportation", "can", "use", "active", "et", "al", "travel", "study", "elsevier", "na", "sciencedirect", "journal", "transport", "mode", "modes", "children", "the", "use", "n", "s", "schools", "also", "p", "b", "used", "number", "data", "research", "variable", "variables", "author", "authors", "researcher", "researchers", "project", "intrument", "may")

corpusNum <- c()

# Iterating through each subfolders 
m <- 1
for (subFolder in subFolders){

  pdfLocation <- subFolder
  if (file.exists(paste(pdfLocation, "txt"))){
    txtLocation <- paste(pdfLocation, "txt")
  }else{
    dir.create(paste(pdfLocation, "txt"))
    txtLocation <- paste(pdfLocation, "txt")
  }
  fileBaseNames <- list.files(pdfLocation, pattern = "*.pdf", full.names=FALSE)
  filePath <- list.files(pdfLocation, pattern = "*.pdf", full.names=TRUE)
  n <- 1

  for (address in filePath)
{
doc <- readPDF(control = list(text = "-layout"))(elem = list(uri = address), language = "en")
test <- doc$content

# "file_path_sans_ext(basename())" from the tools package strips the base name for the file of interest and removed the extension at the end which as seen below allow for the replacement of .pdf for .txt 
# use of 'sep = ""' argument is important to prevent the addition of spaces, which is the default for the paste function 
newFile <- paste(txtLocation, "/", file_path_sans_ext(basename(address)), ".txt", sep = "")
write(test, file = newFile)

####print(paste(file_path_sans_ext(basename(address)), ".txt Has been created", sep = ""))

# Creates uniquely id text store in 'Environment'
assign(paste("text", n, sep = ""), readChar(newFile, file.info(newFile)$size))
n <- n+1
  }

text <- vector()
doc_id <- c(2:n-1)

for (i in seq(1, n-1, 1))
{
text[i] <- eval(parse(text = paste("text", i, sep = "")))
}

textDF <- data.frame(doc_id, text)

temp_df_source <- DataframeSource(textDF)
tempCorpus <- VCorpus(temp_df_source)

# corpus cleaning 
tempCorpus <- tm_map(tempCorpus, stripWhitespace)
tempCorpus <- tm_map(tempCorpus, content_transformer(tolower))
tempCorpus <- tm_map(tempCorpus, removePunctuation)
tempCorpus <- tm_map(tempCorpus, removeWords, stopwords("english"))
tempCorpus <- tm_map(tempCorpus, removeWords, eval(parse(text = paste("NAwords", m, sep = ""))))
tempCorpus <- tm_map(tempCorpus, removeNumbers)

assign(paste("cleanTextDF", m, sep = ""), data.frame(text = sapply(tempCorpus, as.character), stringsAsFactors = FALSE))

assign(paste("corpus", m, sep= ""), tempCorpus)

corpusNum <- c(corpusNum, paste(m))

# removed tempory texts for next iteration
m <- m + 1 
rm(list=ls(pattern = "^text"))

}

corpusKey <- data.frame(corpusNum, subFolders)

rm(temp_df_source, tempCorpus, address, doc_id, fileBaseNames, filePath, i, m, n, newFile, pdfLocation, subFolder, test, txtLocation)
```

Creating Term-Document Matrix and Document-Term Matrix with the fully cleaned corpora.
```{r}
# n - number of corpora
n <- 4
for (i in seq(1, n, 1))
{
tempCorpus <-  eval(parse(text = paste("corpus", i, sep = "")))
TDM <- TermDocumentMatrix(tempCorpus)
DTM <- DocumentTermMatrix(tempCorpus)
matrix1 <- as.matrix(TDM)
matrix2 <- as.matrix(DTM)

assign(paste("corpus", i, "TDM", sep= ""), matrix1)
assign(paste("corpus", i, "DTM", sep= ""), matrix2)

}
rm(matrix1, matrix2)
```

# Vocabulary

```{r}
# n - number of corpora
n <- 4

for (i in seq(1, n, 1))
{
  tempDF <- eval(parse(text = paste("cleanTextDF", i, sep = "")))
  it_train = itoken(tempDF$text, progressbar = FALSE)
  assign(paste("vocab", i, sep= ""), create_vocabulary(it_train))
  tempVocab <- create_vocabulary(it_train)
  assign(paste("vocab", i, sep= ""), prune_vocabulary(tempVocab, term_count_min = 5))
}

rm(tempVocab)
```

# Vector Sort

```{r}
#eval(parse(text = paste("NAwords", 1, sep = "")))
NAwords1

```